---
title: "History of db-benchmark timings"
output:
  html_document:
    fig_width: 9
    fig_height: 6
    self_contained: no
    toc: false
    includes:
      in_header: ga.html
---
```{r render, include=FALSE}
# Rscript -e 'rmarkdown::render("./_report/history.Rmd", output_dir="public")' && xdg-open public/history.html
```

```{r opts, echo=FALSE}
knitr::opts_knit$set(root.dir="..")
knitr::opts_chunk$set(echo=FALSE, cache=FALSE)
```

```{r init, message=FALSE}
source("./_report/report.R", chdir=TRUE)    ## get_report_status_file
source("./_helpers/helpers.R", chdir=TRUE)  ## pretty_sci
library(data.table)
library(lattice)
fctr = function(x, rev=FALSE) {
  levels = unique(x)
  factor(x, levels=if (rev) rev(levels) else levels)
}
.nodename = "mr-dl11"
data = function() {
  d = fread("time.csv")[nodename==.nodename & batch>=1546293600] ## from 2019
  d[run==1L & !is.na(batch)
    ] -> d
  q = fread("_control/questions.csv")
  q[, c("question_group","question") := list(fctr(question_group, rev=TRUE), fctr(question))
    ] -> q
  q[d, on=c("task","question"), nomatch=NULL, list(
    task=i.task, batch=i.batch, solution=fctr(i.solution), data=fctr(substr(i.data, 4L, nchar(i.data))), in_rows=i.in_rows,
    version=i.version, git=i.git,
    time_sec=i.time_sec,
    question_group=x.question_group, question=x.question
    )] -> d
  d[CJ(batch, task, solution, data, question, unique=TRUE), on=c("batch","task","solution","data","question")
    ][, "batch" := as.POSIXct(batch, origin="1970-01-01")
      ][]
}
plot = function(d, s, n, t) {
  tryCatch({
    f = sapply(d, is.factor)
    dd = d[task==t & solution==s & in_rows==n][, which(f) := lapply(.SD, factor), .SDcols=f]
    if (uniqueN(dd$batch)<2L) stop("not enough ticks on X axis") ## workaround deepayan/lattice#19
    xyplot(
      data=dd,
      time_sec~batch|data+question_group,
      groups=question,
      auto.key=list(columns=length(levels(dd$question_group)), lines=TRUE, points=FALSE),
      type="l",
      scales=list(y=list(relation="free", log=TRUE, equispaced.log=FALSE)),
      xlab="benchmark run", ylab="query seconds (log)", main=paste(s, t, pretty_sci(n))
    )
  }, error = function(e) message(e$message))
}
d = data()
```

## Solution {.tabset}

### data.table {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r datatable.groupby.1e7}
plot(d, "data.table", 1e7, "groupby")
```

##### 5 GB

```{r datatable.groupby.1e8}
plot(d, "data.table", 1e8, "groupby")
```

##### 50 GB {.active}

```{r datatable.groupby.1e9}
plot(d, "data.table", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r datatable.join.1e7}
plot(d, "data.table", 1e7, "join")
```

##### 5 GB {.active}

```{r datatable.join.1e8}
plot(d, "data.table", 1e8, "join")
```

##### 50 GB

```{r datatable.join.1e9}
plot(d, "data.table", 1e9, "join")
```

#### groupby2014 {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r datatable.groupby2014.1e7}
plot(d, "data.table", 1e7, "groupby2014")
```

##### 5 GB

```{r datatable.groupby2014.1e8}
plot(d, "data.table", 1e8, "groupby2014")
```

##### 50 GB {.active}

```{r datatable.groupby2014.1e9}
plot(d, "data.table", 1e9, "groupby2014")
```

### pydatatable {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r pydatatable.groupby.1e7}
plot(d, "pydatatable", 1e7, "groupby")
```

##### 5 GB

```{r pydatatable.groupby.1e8}
plot(d, "pydatatable", 1e8, "groupby")
```

##### 50 GB {.active}

```{r pydatatable.groupby.1e9}
plot(d, "pydatatable", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r pydatatable.join.1e7}
plot(d, "pydatatable", 1e7, "join")
```

##### 5 GB {.active}

```{r pydatatable.join.1e8}
plot(d, "pydatatable", 1e8, "join")
```

##### 50 GB

```{r pydatatable.join.1e9}
plot(d, "pydatatable", 1e9, "join")
```

### pandas {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r pandas.groupby.1e7}
plot(d, "pandas", 1e7, "groupby")
```

##### 5 GB

```{r pandas.groupby.1e8}
plot(d, "pandas", 1e8, "groupby")
```

##### 50 GB {.active}

```{r pandas.groupby.1e9}
plot(d, "pandas", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r pandas.join.1e7}
plot(d, "pandas", 1e7, "join")
```

##### 5 GB {.active}

```{r pandas.join.1e8}
plot(d, "pandas", 1e8, "join")
```

##### 50 GB

```{r pandas.join.1e9}
plot(d, "pandas", 1e9, "join")
```

#### groupby2014 {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r pandas.groupby2014.1e7}
plot(d, "pandas", 1e7, "groupby2014")
```

##### 5 GB

```{r pandas.groupby2014.1e8}
plot(d, "pandas", 1e8, "groupby2014")
```

##### 50 GB {.active}

```{r pandas.groupby2014.1e9}
plot(d, "pandas", 1e9, "groupby2014")
```

### dplyr {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r dplyr.groupby.1e7}
plot(d, "dplyr", 1e7, "groupby")
```

##### 5 GB

```{r dplyr.groupby.1e8}
plot(d, "dplyr", 1e8, "groupby")
```

##### 50 GB {.active}

```{r dplyr.groupby.1e9}
plot(d, "dplyr", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r dplyr.join.1e7}
plot(d, "dplyr", 1e7, "join")
```

##### 5 GB {.active}

```{r dplyr.join.1e8}
plot(d, "dplyr", 1e8, "join")
```

##### 50 GB

```{r dplyr.join.1e9}
plot(d, "dplyr", 1e9, "join")
```

#### groupby2014 {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r dplyr.groupby2014.1e7}
plot(d, "dplyr", 1e7, "groupby2014")
```

##### 5 GB

```{r dplyr.groupby2014.1e8}
plot(d, "dplyr", 1e8, "groupby2014")
```

##### 50 GB {.active}

```{r dplyr.groupby2014.1e9}
plot(d, "dplyr", 1e9, "groupby2014")
```

### dask {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r dask.groupby.1e7}
plot(d, "dask", 1e7, "groupby")
```

##### 5 GB

```{r dask.groupby.1e8}
plot(d, "dask", 1e8, "groupby")
```

##### 50 GB {.active}

```{r dask.groupby.1e9}
plot(d, "dask", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r dask.join.1e7}
plot(d, "dask", 1e7, "join")
```

##### 5 GB {.active}

```{r dask.join.1e8}
plot(d, "dask", 1e8, "join")
```

##### 50 GB

```{r dask.join.1e9}
plot(d, "dask", 1e9, "join")
```

### spark {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r spark.groupby.1e7}
plot(d, "spark", 1e7, "groupby")
```

##### 5 GB

```{r spark.groupby.1e8}
plot(d, "spark", 1e8, "groupby")
```

##### 50 GB {.active}

```{r spark.groupby.1e9}
plot(d, "spark", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r spark.join.1e7}
plot(d, "spark", 1e7, "join")
```

##### 5 GB {.active}

```{r spark.join.1e8}
plot(d, "spark", 1e8, "join")
```

##### 50 GB

```{r spark.join.1e9}
plot(d, "spark", 1e9, "join")
```

### juliadf {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r juliadf.groupby.1e7}
plot(d, "juliadf", 1e7, "groupby")
```

##### 5 GB

```{r juliadf.groupby.1e8}
plot(d, "juliadf", 1e8, "groupby")
```

##### 50 GB {.active}

```{r juliadf.groupby.1e9}
plot(d, "juliadf", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r juliadf.join.1e7}
plot(d, "juliadf", 1e7, "join")
```

##### 5 GB {.active}

```{r juliadf.join.1e8}
plot(d, "juliadf", 1e8, "join")
```

##### 50 GB

```{r juliadf.join.1e9}
plot(d, "juliadf", 1e9, "join")
```

### cudf {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r cudf.groupby.1e7}
plot(d, "cudf", 1e7, "groupby")
```

##### 5 GB

```{r cudf.groupby.1e8}
plot(d, "cudf", 1e8, "groupby")
```

##### 50 GB {.active}

```{r cudf.groupby.1e9}
plot(d, "cudf", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r cudf.join.1e7}
plot(d, "cudf", 1e7, "join")
```

##### 5 GB {.active}

```{r cudf.join.1e8}
plot(d, "cudf", 1e8, "join")
```

##### 50 GB

```{r cudf.join.1e9}
plot(d, "cudf", 1e9, "join")
```

### clickhouse {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r clickhouse.groupby.1e7}
plot(d, "clickhouse", 1e7, "groupby")
```

##### 5 GB

```{r clickhouse.groupby.1e8}
plot(d, "clickhouse", 1e8, "groupby")
```

##### 50 GB {.active}

```{r clickhouse.groupby.1e9}
plot(d, "clickhouse", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r clickhouse.join.1e7}
plot(d, "clickhouse", 1e7, "join")
```

##### 5 GB {.active}

```{r clickhouse.join.1e8}
plot(d, "clickhouse", 1e8, "join")
```

##### 50 GB

```{r clickhouse.join.1e9}
plot(d, "clickhouse", 1e9, "join")
```

### polars {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r polars.groupby.1e7}
plot(d, "polars", 1e7, "groupby")
```

##### 5 GB

```{r polars.groupby.1e8}
plot(d, "polars", 1e8, "groupby")
```

##### 50 GB {.active}

```{r polars.groupby.1e9}
plot(d, "polars", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r polars.join.1e7}
plot(d, "polars", 1e7, "join")
```

##### 5 GB {.active}

```{r polars.join.1e8}
plot(d, "polars", 1e8, "join")
```

##### 50 GB

```{r polars.join.1e9}
plot(d, "polars", 1e9, "join")
```

### arrow {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r arrow.groupby.1e7}
plot(d, "arrow", 1e7, "groupby")
```

##### 5 GB

```{r arrow.groupby.1e8}
plot(d, "arrow", 1e8, "groupby")
```

##### 50 GB {.active}

```{r arrow.groupby.1e9}
plot(d, "arrow", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r arrow.join.1e7}
plot(d, "arrow", 1e7, "join")
```

##### 5 GB {.active}

```{r arrow.join.1e8}
plot(d, "arrow", 1e8, "join")
```

##### 50 GB

```{r arrow.join.1e9}
plot(d, "arrow", 1e9, "join")
```


### duckdb {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r duckdb.groupby.1e7}
plot(d, "duckdb", 1e7, "groupby")
```

##### 5 GB

```{r duckdb.groupby.1e8}
plot(d, "duckdb", 1e8, "groupby")
```

##### 50 GB {.active}

```{r duckdb.groupby.1e9}
plot(d, "duckdb", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r duckdb.join.1e7}
plot(d, "duckdb", 1e7, "join")
```

##### 5 GB {.active}

```{r duckdb.join.1e8}
plot(d, "duckdb", 1e8, "join")
```

##### 50 GB

```{r duckdb.join.1e9}
plot(d, "duckdb", 1e9, "join")
```

### duckdb-latest {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r duckdb-latest.groupby.1e7}
plot(d, "duckdb-latest", 1e7, "groupby")
```

##### 5 GB

```{r duckdb-latest.groupby.1e8}
plot(d, "duckdb-latest", 1e8, "groupby")
```

##### 50 GB {.active}

```{r duckdb-latest.groupby.1e9}
plot(d, "duckdb-latest", 1e9, "groupby")
```

#### join {.tabset .tabset-fade .tabset-pills}

##### 0.5 GB

```{r duckdb-latest.join.1e7}
plot(d, "duckdb-latest", 1e7, "join")
```

##### 5 GB {.active}

```{r duckdb-latest.join.1e8}
plot(d, "duckdb-latest", 1e8, "join")
```

##### 50 GB

```{r duckdb-latest.join.1e9}
plot(d, "duckdb-latest", 1e9, "join")
```

## Details

### Environment

Nodename: `r .nodename`

------

Report was generated on: `r format(Sys.time(), usetz=TRUE)`.  

```{r status_set_success}
cat("history\n", file=get_report_status_file(), append=TRUE)
```
