---
title: "History of db-benchmark timings"
output:
  html_document:
    self_contained: no
    toc: false
    includes:
      in_header: ga.html
---
```{r render, include=FALSE}
# Rscript -e 'rmarkdown::render("./_report/history.Rmd", output_dir="public")' && xdg-open public/history.html
```

```{r opts, echo=FALSE}
knitr::opts_knit$set(root.dir="..")
knitr::opts_chunk$set(echo=FALSE, cache=FALSE)
```

```{r init, message=FALSE}
library(data.table)
library(formattable)
library(rpivotTable)
source("./_report/report.R")
```

```{r lib_fmt_tbls}
# format details table
fmt = function(x, caption=NULL) {
  sign_formatter = formatter("span", style = x ~ style(
    color = ifelse(x > 0.05, "red", ifelse(x < -0.05, "green", "black")),
    "font-weight" = ifelse(x > 0.25, "bold", ifelse(x < -0.25, "bold", "normal"))
  ))
  x = copy(x)[, LNd := NULL]
  mett = c("LNt","LRt","LWt","LMt","LYt")
  x[, (mett) := lapply(.SD, formattable, format="f", digits=1), .SDcols=mett]
  metd = c("LRd","LWd","LMd","LYd")
  x[, (metd) := lapply(.SD, percent, format="d"), .SDcols=metd]
  x[, c(mett,metd) := lapply(.SD, prefix, prefix="", na.text=""), .SDcols=c(mett,metd)]
  dims = c(iquestion="Q",in_rows="Size",k="K",na="NA fraction",sort="Sorted")
  metrics = c(LNt="Now", LRt="LastRun", LRd="LR \u2206", LWt="LastWeek", LWd="LW \u2206", LMt="LastMonth", LMd="LM \u2206", LYt="LastYear", LTd="LY \u2206")
  labels = c(dims[match(names(x), names(dims), 0L)], metrics)
  if (length(labels)!=length(x))
    stop(paste0("names(x): ", paste(names(x), collapse=","), "\n"), paste0("  labels: ", paste(labels, collapse=",")))
  formattable(x, list(
    #"q" = if (length(var)) I else color_tile("transparent", "lightpink"),
    #"k" = color_tile("transparent", "lightpink"),
    #"na" = color_tile("transparent", "lightpink"),
    #"sort" = color_tile("transparent", "lightpink"),
    LNt = formatter("span", style = x ~ style("font-weight" = "bold")),
    LRd = sign_formatter, LWd=sign_formatter, LMd=sign_formatter, LYd=sign_formatter
  ), col.names=labels, caption=caption)
}
```

```{r exec_groupby_overview, eval=FALSE}
## re-usable chunk for solution+task filtered data as 'dt' due to
# https://stackoverflow.com/questions/42024921/why-does-kable-not-print-when-used-within-a-function-in-rmarkdown
fmt(
  caption="by Question (k=1e2, no NA, unsorted)",
  dt[k=="1e2" & na=="0" & sort=="0", dcast(.SD, iquestion~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)]
)
fmt(
  caption="by Size (k=1e2, no NA, unsorted)",
  dt[k=="1e2" & na=="0" & sort=="0", dcast(.SD, in_rows~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)]
)
fmt(
  caption="by K cardinality (no NA, unsorted)",
  dt[na=="0" & sort=="0", dcast(.SD, k~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)]
)
fmt(
  caption="by NA fraction (k=1e2, unsorted)",
  dt[k=="1e2" & sort=="0", dcast(.SD, na~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)]
)
fmt(
  caption="by Sortedness (k=1e2, no NA)",
  dt[k=="1e2" & na=="0", dcast(.SD, sort~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)]
)
```

```{r exec_join_overview, eval=FALSE}
fmt(
  caption="by Question (no NA, unsorted)",
  dt[na=="0" & sort=="0", dcast(.SD, iquestion~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)]
)
fmt(
  caption="by Size (no NA, unsorted)",
  dt[na=="0" & sort=="0", dcast(.SD, in_rows~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)]
)
fmt(
  caption="by NA fraction (unsorted)",
  dt[sort=="0", dcast(.SD, na~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)]
)
fmt(
  caption="by Sortedness (no NA)",
  dt[na=="0", dcast(.SD, sort~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)]
)
```

```{r lib_data}
get_batches = function(x) {
  x = unique(x)
  #p = function(x) as.POSIXct(x, origin="1970-01-01")
  secd = 24*60*60
  last = function(x) tail(x, 1L)[1L]
  # recent vs: previous, 1w ago, 1m ago, 1y ago
  b = setNames(rep(NA_integer_, 5L), c("LN","LR","LW","LM","LY"))
  b[["LN"]] = last(x)
  stopifnot(!is.na(b[["LN"]]))
  b[["LY"]] = last(x[x < (b[["LN"]]-365*secd)])
  b[["LM"]] = last(x[between(x, b[["LY"]], b[["LN"]]-30*secd, incbounds=FALSE)])
  b[["LW"]] = last(x[between(x, b[["LM"]], b[["LN"]]-7*secd, incbounds=FALSE)])
  b[["LR"]] = last(x[between(x, b[["LW"]], b[["LN"]], incbounds=FALSE)])
  b
}
do_metrics = function(s, d) {
  d = d[solution==s, .(task, in_rows, k, na, sort, iquestion, time_sec_1, time_sec_2, batch, version, git)]
  tt = as.character(unique(d$task))
  ans = setNames(vector("list", length(tt)), tt)
  for (t in tt) { # t=tt[1]
    # recent vs: previous, 1w ago, 1m ago, 1y ago
    b = get_batches(d[task==t, batch])
    metrics = data.table(metric=factor(names(b), levels=names(b)), batch=b)
    # deal with duplicates when solution does not run for a week, then last run is same as last week
    #metrics = data.table(metric=names(b), batch=b)[, .(metric=tail(metric, 1L)), batch][.(names(b)), on="metric", .(metric=factor(metric, levels=names(b)), batch)]
    non_na_metric = names(b)[1L]
    stopifnot(metrics[metric==non_na_metric, !is.na(batch)])
    dd = d[task==t][metrics[!is.na(batch)], on="batch"]
    if (nrow(dd[is.na(task) | task!=t])) browser()
    stopifnot(nrow(dd[, if (.N > 1L) .SD,, c("task","in_rows","k","na","sort","iquestion","metric")])==0L)
    if (length(missing_metrics <- setdiff(as.character(metrics$metric), as.character(unique(dd$metric))))) { ## fill empty to match number of columns always
      metric_fill = function(missing_metric, dd, non_na_metric) {
        missing_dd = dd[metric==non_na_metric][, `:=`(metric=missing_metric, time_sec_1=NA_real_, time_sec_2=NA_real_, batch=NA_integer_, version=NA_integer_, git=NA_integer_)]
        stopifnot(nrow(missing_dd[, if (.N > 1L) .SD,, c("task","in_rows","k","na","sort","iquestion","metric")])==0L)
        if (nrow(missing_dd[is.na(task) | task!=t])) browser()
        missing_dd
      }
      dd = rbindlist(c(list(dd), lapply(missing_metrics, metric_fill, dd, non_na_metric)))
      setorderv(dd, "metric")
    }
    stopifnot(nrow(dd[, if (.N > 1L) .SD,, c("task","in_rows","k","na","sort","iquestion","metric")])==0L)
    if (nrow(dd[is.na(task) | task!=t])) browser()
    ans[[t]] = dd
  }
  rbindlist(ans)
}
do_delta = function(dd) {
  dd = copy(dd)
  setnames(dd, c("time_sec_1","time_sec_2"), c("run_1","run_2"))
  now = dd[metric=="LN", .(solution, task, in_rows, k, na, sort, iquestion, run_1, run_2)]
  stopifnot(nrow(now[, if (.N>1L) .SD,, c("solution","task","in_rows","k","na","sort","iquestion")])==0L,
            nrow(dd[is.na(metric)])==0L)
  delta = dd[now, on=c("solution","task","in_rows","k","na","sort","iquestion"),
             .(solution, task, in_rows, k, na, sort, iquestion, run_1=(i.run_1-x.run_1)/x.run_1, run_2=(i.run_2-x.run_2)/x.run_2, batch, version, git, metric=x.metric)]
  stopifnot(nrow(delta[is.na(metric)])==0L, is.factor(dd$metric), is.factor(delta$metric))
  #return(list(dd=dd, delta=delta))
  levels(delta$metric) = paste0(levels(dd$metric), "d")
  levels(dd$metric) = paste0(levels(dd$metric), "t")
  ddd = rbindlist(list(dd, delta))
  if (nrow(ddd[is.na(metric)])) browser()
  ddd[, "metric" := factor(metric, levels=c("LNt","LNd","LRt","LRd","LWt","LWd","LMt","LMd","LYt","LYd"))] # reorder levels
  if (nrow(ddd[is.na(metric)])) browser()
  ddd
}
```

Work in progress: performance regression tracking

Note:

- missing `LastRun` is likely to occur when no runs happened for more than a week, so the values would be the same as `LastWeek`.
- missing values are likely to occur due to the missing timing values, for example, due to the performance regression a question haven't finished, then we should not skip it during calculating `mean`.
- first run timings are used in calculations.
- timing metrics are not normalized when passing to `mean`, thus `mean` over data sizes (1e7, 1e8, 1e9)  will do something like `mean(c(0.1, 1, 10))`.

```{r data}
# load data
ld = time_logs()[nodename==nodename[.N]]

# for each solution find 5 runs: now, previous, 1w ago, 1m ago, 1 year ago
dd = rbindlist(lapply(setNames(nm=levels(ld$solution)), do_metrics, ld), idcol="solution")

# do 'delta' metrics
ddd = do_delta(dd)
```

## Solution {.tabset}

### data.table {.tabset .tabset-fade .tabset-pills}

#### groupby {.tabset .tabset-dropdown}

```{r datatable.groupby}
dt = ddd[solution=="data.table" & task=="groupby"]
```

##### Overview

```{r datatable.groupby.overview, ref.label="exec_groupby_overview"}
```

##### Question

```{r datatable.groupby.q}
fmt(dt[k=="1e2" & na=="0" & sort=="0", dcast(.SD, iquestion+in_rows~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)],
    caption="by Question Size (k=1e2, no NA, unsorted)")
fmt(dt[na=="0" & sort=="0", dcast(.SD, iquestion+k~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)],
    caption="by Question K cardinality (no NA, unsorted)")
```

##### K cardinality

```{r datatable.groupby.k}
fmt(dt[na=="0" & sort=="0", dcast(.SD, k+in_rows~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)],
    caption="by K cardinality Size (no NA, unsorted)")
fmt(dt[na=="0" & sort=="0", dcast(.SD, k+iquestion~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)],
    caption="by K cardinality Question (no NA, unsorted)")
```

##### NA fraction

```{r datatable.groupby.na}
fmt(dt[k=="1e2" & sort=="0", dcast(.SD, na+in_rows~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)],
    caption="by NA fraction Size (k=1e2, unsorted)")
```

##### sortedness

```{r datatable.groupby.sort}
fmt(dt[k=="1e2" & na=="0", dcast(.SD, sort+in_rows~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)],
    caption="by NA fraction Size (k=1e2, no NA)")
```

#### join {.tabset .tabset-dropdown}

```{r datatable.join}
dt = ddd[solution=="data.table" & task=="join"]
```

##### Overview

```{r datatable.join.overview, ref.label="exec_join_overview"}
```

##### Question

```{r datatable.join.q}
fmt(dt[na=="0" & sort=="0", dcast(.SD, iquestion+in_rows~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)],
    caption="by Question Size (no NA, unsorted)")
```

##### NA fraction

```{r datatable.join.na}
fmt(dt[sort=="0", dcast(.SD, na+in_rows~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)],
    caption="by NA fraction Size (unsorted)")
```

##### sortedness

```{r datatable.join.sort}
fmt(dt[na=="0", dcast(.SD, sort+in_rows~metric, value.var="run_1", fun.aggregate=mean, na.rm=FALSE)],
    caption="by NA fraction Size (no NA)")
```

### dplyr

### pandas

### pydatatable

### spark

### cudf

### dask

### juliadf

### clickhouse

## Details {.tabset .tabset-fade .tabset-pills}

### Environment

Nodename: `r ld[1L][["nodename"]]`

### Timings

```{r pivot}
sdcols = c("solution","task","question_group","question","iquestion","data","in_rows","k","na","sorted","time_sec_1","time_sec_2","version","git","chk_time_sec_1","na_time_sec","out_rows","out_cols")
data = ld[, .SD, .SDcols=sdcols]
rpivotTable::rpivotTable(
  data,
  rows = c("task","in_rows","k","sorted","question"),
  cols = "solution",
  aggregatorName = "Average",
  vals = "time_sec_1",
  height = "100%",
  unusedAttrsVertical = TRUE
)
```

## {-}

------

Report was generated on: `r format(Sys.time(), usetz=TRUE)`.  

```{r status_set_success}
cat("history\n", file=get_report_status_file(), append=TRUE)
```
